---
title: "mini-project 2"
author: "Sha Liu"
date: "11/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE, fig.height=4, fig.width=8)
#install.packages("tidyverse")
library(readxl)
library(MASS)
library(arm)
library(mgcv)
library(tidyverse)
library(ggthemes)
library(tidyverse)
library(data.table)
```

```{r}
# covid <- read.csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv",
#                      stringsAsFactors = FALSE) %>%
#   filter(date == max(date) & county != "New York City") 
# 
# kolko <- read_excel("kolko_covid_shareable.xlsx")
# 
# covid <- left_join(covid, kolko, by = c("fips" = "county")) %>%
#   select(-pcases) %>%
#   filter(!is.na(density))
# 
# covid <- rename(covid, cases = cases.x, deaths = deaths.x)
# 
# # Standardized cases by population per county, created a new variable named casespercapita
# covid <- covid %>% 
#   mutate(casespercapita = cases/popestimate2019)

#write.csv(covid, "covid.csv")
```

```{r}
covid <- fread("covid.csv")
```


```{r}
cor(covid$density, covid$lncbsapop)
```

## **Executive summary**					
We explore the relationship between covid confirmed cases and variables like Tract-weighted household density and Natural log of the Core Based Statistical Area population and other variables from January 2020 to present using data from the Github website (https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv). In our data exploratory analysis, we find that ...

## **Relationship between density and COVID incidence**
We begin by examining the relationship between between density and COVID incidence, we standardized cases by population. We found that as the density goes up till around 8 (in log scale), the confirmed cases rate per county goes up. The the confirmed cases per capita slowly goes down, then the density goes up.

```{r}
ggplot(covid, aes(x= density, y = casespercapita)) +
  geom_point() + geom_smooth(se =FALSE) +
  labs(y="Confirmed COVID-19 cases per capita",
       x="Tract-weighted household density",
       title= paste("Figure 1: Confirmed COVID-19 cases per capita as of", max(covid$date)),
       caption = "Data from https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv") +
  theme_minimal() +
  scale_color_colorblind() +
  #scale_y_continuous(labels = scales::percent) + geom_hline(yintercept = 0.025, color = "red") +
  scale_x_log10(labels = scales::comma)
```

## **Relationship between CBSA population size and COVID incidence**
We then examine the relationship between between CBSA population size and COVID incidence. we found that as the Natural log of the Core Based Statistical Area population goes up till 9, the confirmed cases per capita goes up. Then the confirmed cases per capita slowly goes down as the popultion size increases.

```{r}
ggplot(covid, aes(x = lncbsapop, y = casespercapita)) +
  geom_point() + geom_smooth(se = FALSE) +
  labs(y="Confirmed COVID cases per capita",
       x="Natural log of the Core Based Statistical Area population",
       title= paste("Figure 1: Confirmed COVID-19 cases per capita as of", max(covid$date)),
       caption = "Data from https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv") +
  theme_minimal() +
  scale_color_colorblind() +
  scale_y_continuous(labels = scales::percent)
```
## **GDP per capita in 1960 and growth rate since 1960** 

We continue to examine the relationship between COVID incidence with both density and CPBA population size as the only two explanatory variables. We start with faceting by have found that model gam give us a good fit with the data. We use ...

```{r}
#ggplot(covid, aes(x = density, y = casespercapita)) + geom_point(size = 0.7) +
#  facet_grid(~cut_number(lncbsapop, n=6)) + geom_smooth(method = "loess", color = "orange") + #geom_smooth(method = "gam", color = "green") + 
#  labs(y="Confirmed COVID cases rates per County",
#       x="Tract-weighted household density",
#       title="Figure 1: Confirmed COVID cases rates per County as of date = max(covid$date) ",
#       subtitle = "faceted by CBSA population size.",
#       caption = "Data from https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv") +
#  theme_minimal() +
#  scale_color_colorblind() +
#  scale_y_continuous(labels = scales::percent) + 
#  scale_x_log10(labels = scales::comma)

```

## **Interaction**	

I will need an interaction between density and lncbsapop because the correlations between these two is really high. Since there is an interaction, we decide to fit with a non-parametric gam model.

# GAM

# Top 10 correlated varibales

```{r}
# install.packages("pkgconfig")
# devtools::install_github("laresbernardo/lares")
library(lares)
corr_var(covid, # name of dataset
  casespercapita, # name of variable to focus on
  top = 10 # display top 10 correlations
) 
```

```{r}

covid.gam1 = gam(casespercapita ~ s(density, lncbsapop) + s(age60plus) + s(black_pct) + s(seasonal_units) + s(obesity), data = covid, method = "REML")

covid.gam.df1 = data.frame(covid, .fitted = fitted.values(covid.gam1), 
                          .resid = residuals(covid.gam1))

ggplot(covid.gam.df1, aes(x = lncbsapop, y = .resid)) + geom_point() +
  ggtitle("GAM residuals") + scale_x_log10() +
  geom_smooth(method = "lm", se = FALSE)
```


```{r}
ggplot(covid.gam.df1, aes(x = lncbsapop, y = .fitted)) +
  ggtitle("GAM fitted values") +
  geom_point(data = covid, aes(y = casespercapita), size = 0.5) +
  geom_smooth(method = "lm", se = FALSE)


ggplot(covid.gam.df1, aes(x = lncbsapop, y = .fitted)) +
  facet_wrap(~cut_number(log(density), n=3), scales = "free") + ggtitle("GAM fitted values") +
  geom_point(data = covid, aes(y = casespercapita), size = 0.5) + 
  geom_smooth(method = "lm", se = FALSE)

```


# Plot the model

```{r}
# summary(covid.gam)
# covid.grid = expand.grid(lncbsapop = c(5, 7, 9, 11, 13, 15, 17),
#                           density = 0:15523)
# covid.pred = predict(covid.gam, newdata = covid.grid)
# covid.plot.df = data.frame(covid.grid, casespercapita = as.vector(covid.pred))
```


# GLM model

```{r}
glm_covid7 <- glm(cases ~ density * lncbsapop + age60plus + black_pct + seasonal_units + obesity + college + crowded, data = covid, family = quasipoisson, offset = log(popestimate2019))

summary(glm_covid7)

covid_df <- data.frame(covid, 
                       .fitted = fitted.values(glm_covid7),
                       .resid = residuals(glm_covid7))

ggplot(covid_df, aes(density, .resid)) + 
  geom_point() +
  geom_smooth(method = "lm", se=FALSE) +
  scale_x_log10()
```


# 4. ...

```{r}
library(performance)

test <- covid %>% 
  filter(!is.na(density)) %>% 
  filter(date %in% c(as.Date("2020-04-30"), as.Date("2020-08-31"), max(date))) %>% 
  pivot_wider(-contains(c("death", "case", "V1","date")), names_from = date, values_from = cases.x) %>% 
  mutate(First  = `2020-04-30`,
         Second = `2020-08-31` - `2020-04-30`,
         Third  = `2020-11-09` - `2020-08-31`) %>% 
  filter(Third > 0)
  #pivot_longer(First:Third, names_to = "Period", values_to="Period_cases") %>% 
  #group_by(Period) %>% 
  

ggplot(test) +
  geom_smooth(method="lm",aes(density, First)) +
  geom_smooth(method="lm",aes(density, Second),color="red") +
  geom_smooth(method="lm",aes(density, Third),color="yellow") +
  scale_x_log10() +
  scale_y_log10()

#glm_covid1 <- glm(First ~ density, data = test, family = quasipoisson)
#glm_covid2 <- glm(Period_cases ~ density * lncbsapop, data = test)
#glm_covid3 <- glm(Period_cases ~ density * lncbsapop * age60plus, data = test)
#glm_covid4 <- glm(Period_cases ~ density * lncbsapop


#glm_covid5 <- glm(Period_cases ~ density * lncbsapop * age60plus * china_born * obesity, data = test)
#glm_covid6 <- glm(Period_cases ~ density + lncbsapop * age60plus * china_born * obesity, data = test)

glm_covid7 <- glm(First ~ density * lncbsapop + age60plus + black_pct + seasonal_units + obesity + college + crowded, data = test %>% filter(!is.na(First)), family = quasipoisson, offset = log(popestimate2019))
glm_covid8 <- glm(Second ~ density * lncbsapop + age60plus + black_pct + seasonal_units + obesity + college + crowded, data = test %>% filter(!is.na(Second)), family = quasipoisson, offset = log(popestimate2019))
glm_covid9 <- glm(Third ~ density * lncbsapop + age60plus + black_pct + seasonal_units + obesity + college + crowded, data = test %>% filter(!is.na(Third)), family = quasipoisson, offset = log(popestimate2019))

ggplot(glm_covid7, aes(density,residuals(glm_covid7))) +
  geom_point() +
  geom_smooth(method="lm") +
  scale_x_log10()
ggplot(glm_covid8, aes(density,residuals(glm_covid8))) +
  geom_point() +
  geom_smooth(method="lm") +
  scale_x_log10()
ggplot(glm_covid9, aes(density,residuals(glm_covid9))) +
  geom_point() +
  geom_smooth(method="lm") +
  scale_x_log10()

check_model(glm_covid7)
check_model(glm_covid8)
check_model(glm_covid9)

check_distribution(glm_covid7)
check_distribution(glm_covid8)
check_distribution(glm_covid9)

```





